---
title: "Learning Curves"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    social: menu
    source_code: https://github.com/bradleyboehmke/learningCurve-shiny-app
    theme: simplex
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(learningCurve)
library(ggplot2)
library(plotly)
library(DT)
```


Sidebar {.sidebar}
======================================================================

```{r}

# Define inputs

textInput("unit_a", "Starting Unit", 1)
textInput("time_a", "Starting Unit Time", 100)
textInput("unit_b", "Ending Unit", 250)

sliderInput("rate", "Learning Curve Rate:",
    min = 0, max = 1, value = .8
  )

radioButtons("model", "Model:",
             c("Crawford's Unit Learning Curve Function" = "u",
               "Wright's Cumulative Average Learning Curve Function" = "ca"))

#actionButton("goButton", "Compute!")

```

<br>

Use the __Explore Learning Curves__ tab to explore how differing inputs such as starting unit time, learning rate, and type of learning curve function (Crawford's unit function versus Wright's cummulative average function) influences the time (or cost) predicted for the ending unit. Scroll along the unit- and cummulative-level plots to examine predicted time (or cost) for different levels of units produced.  

To learn more click the __About__ tab.  

Application author: [Bradley Boehmke](http://bradleyboehmke.github.io/)


Explore Learning Curves
======================================================================

Row
-----------------------------------------------------------------------

### Unit-level Estimates

```{r}

df <- reactive({
  if(input$model == "u"){
    
  data.frame(Unit = as.numeric(input$unit_a):as.numeric(input$unit_b),
             Time = unit_curve(t = as.numeric(input$time_a),
                               m = as.numeric(input$unit_a),
                               n = as.numeric(input$unit_a):as.numeric(input$unit_b),
                               r = as.numeric(input$rate)),
             Cummulative = cumsum(unit_curve(t = as.numeric(input$time_a),
                                             m = as.numeric(input$unit_a),
                                             n = as.numeric(input$unit_a):as.numeric(input$unit_b),
                                             r = as.numeric(input$rate))))
  } else {
    
    data.frame(Unit = as.numeric(input$unit_a):as.numeric(input$unit_b),
             Time = ca_unit(t = as.numeric(input$time_a),
                            m = as.numeric(input$unit_a),
                            n = as.numeric(input$unit_a):as.numeric(input$unit_b),
                            r = as.numeric(input$rate)),
             Cummulative = cumsum(ca_unit(t = as.numeric(input$time_a), 
                                          m = as.numeric(input$unit_a),
                                          n = as.numeric(input$unit_a):as.numeric(input$unit_b), 
                                          r = as.numeric(input$rate))))
  }
})
```


```{r}

renderPlotly({
  p1 <- ggplot(df(), aes(x = Unit, y = Time)) +
    geom_line() +
    scale_x_continuous("Units Produced", labels = scales::comma) +
    scale_y_continuous("Per Unit Time (or Cost)", labels = scales::comma)
  
  ggplotly(p1)
})
```

### Cummulative-level Estimates

```{r}

renderPlotly({
  p1 <- ggplot(df(), aes(x = Unit, y = Cummulative)) +
    geom_line() +
    scale_x_continuous("Units Produced", labels = scales::comma) +
    scale_y_continuous("Cummulative Time (or Cost)", labels = scales::comma)
})
```

Row
-----------------------------------------------------------------------

### Model Definition

Crawford's unit learning curve model:

$$y_n = t_m * \left(\frac{n}{m}\right)^b$$

Wright's cummulative average learning curve model:

$$y_{n} = t_{m} * \frac{n^{1+b} - (n-1)^{1+b}}{m^{1+b} - (m-1)^{1+b}}$$

both where

&nbsp;&nbsp;&nbsp;&nbsp; $y_n$ = the time (or cost) required for the nth unit of production <br>
&nbsp;&nbsp;&nbsp;&nbsp; $t_{m}$ = time (or cost) required for the mth unit of production <br> 
&nbsp;&nbsp;&nbsp;&nbsp; *m* = mth unit of production <br>
&nbsp;&nbsp;&nbsp;&nbsp; *n* = nth unit you wish to predict the time (or cost) for <br>
&nbsp;&nbsp;&nbsp;&nbsp; *b* = natural slope of the learning curve rate <br>


To compute *b* for a given learning rate:
  
$$b = \frac{log(r)}{log(2)}$$





### Underlying Data

```{r}
df.table <- reactive({
  if(input$model == "u"){
    
  data.frame(Unit = as.numeric(input$unit_a):as.numeric(input$unit_b),
             `Unit Time` = round(unit_curve(t = as.numeric(input$time_a),
                               m = as.numeric(input$unit_a),
                               n = as.numeric(input$unit_a):as.numeric(input$unit_b),
                               r = as.numeric(input$rate)), 0),
             `Cummulative Time` = round(cumsum(unit_curve(t = as.numeric(input$time_a),
                                             m = as.numeric(input$unit_a),
                                             n = as.numeric(input$unit_a):as.numeric(input$unit_b),
                                             r = as.numeric(input$rate))), 0))
  } else {
    
    data.frame(Unit = as.numeric(input$unit_a):as.numeric(input$unit_b),
             `Unit Time` = round(ca_unit(t = as.numeric(input$time_a),
                            m = as.numeric(input$unit_a),
                            n = as.numeric(input$unit_a):as.numeric(input$unit_b),
                            r = as.numeric(input$rate)), 0),
             `Cummulative Time` = round(cumsum(ca_unit(t = as.numeric(input$time_a), 
                                          m = as.numeric(input$unit_a),
                                          n = as.numeric(input$unit_a):as.numeric(input$unit_b), 
                                          r = as.numeric(input$rate))), 0))
  }
})

renderDataTable({
  datatable(df.table(), rownames = FALSE,
            colnames = c('Unit Number', 'Unit Time', 'Cummulative Time'),
            options = list(pageLength = 8, dom = 'tip'))
})
```







About
======================================================================

<h2> Learning Curves </h2>

This application is for the purpose of aiding student comprehension of learning curves by allowing them to explore learning curves interactively. Learning curves are steep in history and have several alternate names such as improvement curves ([Fauber, 1989](http://search.proquest.com/openview/4192c8b3f2e14c8a64ac5c118883aede/1.pdf?pq-origsite=gscholar&cbl=36911)), efficiency curves ([Biørn, 1998](http://link.springer.com/article/10.1007/BF01205997)), experience curves ([Desroches et al., 2013](http://www.sciencedirect.com/science/article/pii/S0301421512008488)), progress curves ([Dutton & Thomas, 1984](https://www.jstor.org/stable/258437?seq=1#page_scan_tab_contents)) and cost improvement curves ([Miller et al., 2012](http://www.tandfonline.com/doi/abs/10.1080/1941658X.2012.682943)). The "learning effect" was first noted by T.P. Wright in the 1920s in connection with aircraft production ([Wright, 1936](http://arc.aiaa.org/doi/abs/10.2514/8.155)). Its use was amplified by experience in connection with aircraft production in WW II. Initially, it was thought to be solely due to the learning of the workers as they repeated their tasks. Later, it was observed that other factors probably entered in, such as improved tools and working conditions, and various management initiatives. Regardless of the exact, or more likely combined, phenomenon we can group these factors together under the general heading of “learning.”

The underlying notion behind learning curves is that when people individually or collectively repeat an activity, there tends to be a gain in efficiency. Generally, this takes the form of a decrease in the time needed to do the activity. Because cost is generally related to time or labor hours consumed, learning curves are very important in industrial cost analysis. A key idea underlying the theory is that every time the production quantity doubles, we can expect a more or less fixed percentage decrease in the effort required to build a single unit (the Crawford theory), or in the average time required to build a group of units (the Wright theory). These decreases occur not in big jumps, but more or less smoothly as production continues.
  
Consequently, mathematical models are used to represent learning curves by computing the efficiencies gained when an activity is repeated.  The two most common models are:

Crawford's unit learning curve model:

$$y_n = t_m * \left(\frac{n}{m}\right)^b$$

Wright's cummulative average learning curve model:

$$y_{n} = t_{m} * \frac{n^{1+b} - (n-1)^{1+b}}{m^{1+b} - (m-1)^{1+b}}$$

both where

&nbsp;&nbsp;&nbsp;&nbsp; $y_n$ = the time (or cost) required for the nth unit of production <br>
&nbsp;&nbsp;&nbsp;&nbsp; $t_{m}$ = time (or cost) required for the mth unit of production <br> 
&nbsp;&nbsp;&nbsp;&nbsp; *m* = mth unit of production <br>
&nbsp;&nbsp;&nbsp;&nbsp; *n* = nth unit you wish to predict the time (or cost) for <br>
&nbsp;&nbsp;&nbsp;&nbsp; *b* = natural slope of the learning curve rate <br>


To compute *b* for a given learning rate (r):
  
$$b = \frac{log(r)}{log(2)}$$

<h3> Colophon </h3>

This app was made by [Brad Boehmke](http://bradleyboehmke.github.io). It leverages the [learningCurve package](https://cran.r-project.org/web/packages/learningCurve/index.html), which has implemented Crawford's and Wright's learning curve functions among others. The application was built with the [Shiny](http://shiny.rstudio.com) framework for the [R programming language](https://www.r-project.org/). The layout was produced with the [flexdashboard](http://rstudio.github.io/flexdashboard/index.html) package, and the charts and tables use [ggplot2](http://ggplot2.org/), [Plotly](http://plot.ly), and [DataTables](https://rstudio.github.io/DT/), all accessed through their corresponding R packages.  Code is available on [GitHub](https://github.com/bradleyboehmke/learningCurve-shiny-app) licensed MIT. Suggestions and criticism welcome: [bradleyboehmke@gmail.com](mailto:bradleyboehmke.com).

<br>
<br>

